A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching

### 2016-10-18 JTB_supp/JTB_supp.Rmd
updates

---
title: "JTB supplement"
author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
date: "October 14, 2016"
output: html_document
bibliography: library.bib
---

@@

This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.

## Parameter values

All parameter values used in the model are listed here along with the relevant literature that provides a basis for the value. In most cases, the units of the values reported in the literature needed to be converted to units appropriate for the model, which was accomplished using a set of stoichiometric conversion factors also based on reported literature (see "R/conversions.R" in this paper's data repository). The specific calculations used to convert values from units reported in the literature to relevant units for the model can be found in the R Markdown code used to produce this document.

@@

    + The Redfield ratio (following @Muller:2009io)

@@

    + Assuming the coral feeds primarily on zooplankton, a general heterotroph has a N:C ratio of 0.2 [@Muller:2009io]

@@

    + @Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day. Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching (see below), giving a value of:

@@

# Import data from Marcelino et al. 2013
md <- read.csv("data/marcelino2013_fig2a.csv")
md$S.H <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
plot(md$A ~ md$S.H, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="A", ylim=c(1,3.2),
     xlab="S:H biomass")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.H), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.H=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit and add to plot
text(0.13,3.1, labels=paste0(c("A = ", as.character(formula(mod))[3]), collapse=""), adj=0)
coefs <- round(coef(mod), 2)
text(0.13,2.9, labels=paste0(c("x=",coefs[1],"; y=",coefs[2],"; z=",coefs[3]), collapse=""), adj=0)
```

## Figure S2. Photosynthesis-Irradiance curves of modeled symbiont.
```{r run1_PI, cache=T, fig.height=3, fig.width=5, echo=F, fig.cap="The amount of light energy quenched through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line). These simulations represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value). The structure of the photosynthesis SU, with the negative feedback between excess light and photosynthesis rate, produces the classic shape of a P-I curve demonstrating photoinhibition, suggesting the mechanism is robust."}
sym_PI(pars=def_pars())
```

>>>>

---
title: "A simplified dynamic bioenergetic model for coral-algal symbiosis and coral bleaching"
author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger M. Nisbet"
output: html_document
bibliography: library.bib
---

# Supplementary Information
This supplement provides justification and derivations from the literature for all parameter values used in the model, as well as additional figures that demonstrate the structure and behavior of the model and its specific components.

@@

## Parameter value derivations
All parameters in the model are listed here along with the relevant literature that provides a basis for the value used. In most cases, the units of the values reported in the literature needed to be converted to units appropriate for the model, which was accomplished using a consistent set of stoichiometric conversion factors also based on reported literature (see "R/conversions.R" in this paper's data repository). The specific calculations used to convert values from the units reported in the literature to the relevant units for the model can be found in the RMarkdown code used to produce this document.

@@

    + This value should approximate the Redfield ratio (following @Muller:2009io).

@@

    + Assuming the coral feeds primarily on zooplankton, a general heterotroph has a N:C ratio of 0.2 [@Muller:2009io].

@@

    + @Gorbunov:2001p4181 report that the capacity of non-photochemical quenching is approximately 4 times the capacity of photochemistry on a sunny day. Therefore, we take the maximum relative capacity of NPQ to be 4 times the maximum rate of photochemical quenching ($j_{CPm}$; see below), giving a value of:

@@

# Import data from Marcelino et al. 2013
md <- read.csv("data/marcelino2013_fig2a.csv")
md$S.H <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
par(mfrow=c(1,1), mar=c(3,3,3,3), mgp=c(1.2,0,0), cex=1, tck=0.025)
plot(md$A ~ md$S.H, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="A (Amplification factor)", ylim=c(1,3.2),
     xlab="S:H biomass")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.H), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.H=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit and add to plot
text(0.13,3.1, labels=paste0(c("A = ", as.character(formula(mod))[3]), collapse=""), adj=0)
coefs <- round(coef(mod), 2)
text(0.13,2.9, labels=paste0(c("x=",coefs[1],"; y=",coefs[2],"; z=",coefs[3]), collapse=""), adj=0)
```

## Figure S2. Photosynthesis-Irradiance curves of modeled symbiont.
```{r run1_PI, cache=F, fig.height=4, fig.width=7, echo=F, fig.cap="The amount of light energy quenched through photochemistry (red line) and non-photochemical quenching (black line), and the relative production of reactive oxygen species (orange line) are plotted as a function of the scalar irradiance experienced by the symbiont (note different y-axes for each line). These simulations represent the behavior of the photosynthesis SU independent of the host and without CO~2~ limitation of photosynthesis (*j~CO2~* set to a very high value). The structure of the photosynthesis SU, with the negative feedback between excess light and photosynthesis rate, produces the classic shape of a P-I curve demonstrating photoinhibition, suggesting the mechanism is robust."}
sym_PI(pars=def_pars())
```
